Evolution of Sex-Specific Recombination Rate

### 2020-05-19 doc/MS_v2.Rmd
start integrating Bret's new edits

@@

```{r Fig3.example.double.triangle, echo=FALSE}
#load an example two triangle plot

single.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/single_chrm_yellow.png')

@@

```{r Q1.ggplot_IFD.triangle, echo=FALSE}
#IFD.Male.DOM.top.tri.plot <- ggplot( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ),
#  aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
#  geom_jitter(alpha=.4)+
#  ggtitle("Normalized IFD\nDom")+
#  scale_color_manual(values=colors_of_strains)+xlim(c(0,1))+ylim(c(0,1))+
#  theme(legend.position="none")+theme_classic()+geom_abline(intercept = 0, slope = 1)

@@

#this works
top.triangle2 <- function(my.y, my.x, my.data) {
  plot(my.y ~ my.x, data= my.data, axes=FALSE, ylab='', xlab="", 
       main='', xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1,
       col =  colors_of_strains[current.strain])

@@

bottom.triangle2 <- function(my.y, my.x, my.data){
  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
          col =  colors_of_strains[current.strain])

@@

@@

```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE}

@@

```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}

@@

```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE, fig.height=3, fig.width=8, fig.cap="Variation in total SC area per cell across strains and Relationship between total sc and mean MLH1 count per cell across male mice. A)	 Mean total SC area per mouse across subspecies and recombination groups. High- and low-recombining groups are indicated by 1 and 0 respectively. (Horizontal lines indicate strain averages for total SC area.)B) Mouse averaged rates of MLH1 and total SC per cell"}

@@

Q2.show.nrmF1

@@

The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = r Q2.1COpos_mus.av_Musc_ttest$p.value and p = r Q2.1COpos_mus.av_Mol_ttest$p.value for _musculus_ and _molossinus_, respectively). While there are significant strain effects for _domesticus^WSB^_ and _molossinus^MOLF^_ in a general linear model with normalized position as the dependent variable, this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate since these two strains do not vary significantly from the male-specific rate.

@@

@@

  mtext("Position.2.Male", 1, 1.5, at= mean(par()$usr[1:2])) #at should be changing the position
  #above is just changing the text not the numbers
  #first number, on which side of the plot (1=bottom, 2=left, 3=top, 4=right)

@@

  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +4)

  x.at <- axisTicks(par()$usr[1:2], 0) #where the ticks should go
  axis(side=1, las=1, pos=0, at=x.at, 
       labels=F, xpd=TRUE)
#las, orientation, at location of line
  #labels automatic labels
  #pos
  #xpd - clipping

   mtext(seq(0, 1, .2), 1, 1, at=x.at)
  #second / third val negative moves the valu up

  axis(4, las=1, pos=1) #this is the y axis
 #   axis(4, las=1, pos=1 + x.dist)
  #this is calling axis twice, not sure why
  lines(0:1, 0:1, xpd=TRUE) #diagonal line
}

@@

  #adjusting this mar lowers the margins (the edges)
  op <- par(mar=c(1.5, 1.5, 1.5, 1.5), oma=c(2, 0, 0, 2))

  #   op <- par(mar=c(1.5, 2, 4, 6), oma=c(2, 0, 0, 2))
  Q2.bottom.triangle2(my.y.bottom, my.x.bottom, my.data) #x.dist

@@

@@

Q2.plot1.WSB <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_WSB.png')
Q2.plot2.LEW <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_LEW.png')
Q2.plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_G.png')

Q2.plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_KAZ.png')
Q2.plot5.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_SKIVE.png')
Q2.plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_PWD.png')

Q2.plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MSM.png')
Q2.plot8.MOLF <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MOLF.png')

grid.arrange(rasterGrob(Q2.plot1.WSB), rasterGrob(Q2.plot2.LEW), rasterGrob(Q2.plot3.G),
             rasterGrob(Q2.plot4.KAZ), rasterGrob(Q2.plot5.SKIVE), rasterGrob(Q2.plot6.PWD),
             rasterGrob(Q2.plot7.MSM), rasterGrob(Q2.plot8.MOLF), 
             ncol=3)

@@

@@

>>>>

<v2 - Bret's edits, v1>

@@

@@

# IFD double triangle

# Figure 1

@@

#this works
top.triangle <- function(my.y, my.x, my.data) {
  plot(my.y ~ my.x, data= my.data, axes=FALSE, ylab='', xlab="", 
       main='', xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1,
       col =  colors_of_strains[current.strain])

@@

bottom.triangle <- function(my.y, my.x, my.data){
  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
          col =  colors_of_strains[current.strain])

@@

  par(op)
  RP[[i]] <- recordPlot()
  dev.off()
}
```

@@

```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}

@@

```{r Q2.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}

@@

```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE, fig.height=3, fig.width=8, fig.cap="Variation in total SC area per cell across strains and Relationship between total sc and mean MLH1 count per cell across male mice. A)	 Mean total SC area per mouse across subspecies and recombination groups. High- and low-recombining groups are indicated by 1 and 0 respectively. (Horizontal lines indicate strain averages for total SC area.)B) Mouse averaged rates of MLH1 and total SC per cell", echo=FALSE}

@@

@@

The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = r Q2.1COpos_mus.av_Musc_ttest$p.value and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value for _musculus_ and _molossinus_, respectively). While there are significant strain effects for _domesticus^WSB^_ and _molossinus^MOLF^_ in a general linear model with normalized position as the dependent variable, this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate since these two strains do not vary significantly from the male-specific rate.

@@

# rewrite double triangle

@@

  mtext("Position.2.Male", 1, -2, at= mean(par()$usr[1:2]) + 0)

  #the second val will move the x lable when negative

  #at should be changing the position
  #above is just changing the text not the numbers
  #first number, on which side of the plot (1=bottom, 2=left, 3=top, 4=right)

@@

  mtext("Position.1.Male", 4, 0, padj=par()$usr[1] + .5) #last value at .5 is good for moving y label

  x.at <- axisTicks(par()$usr[1:2], 0) #where the ticks should go

  axis(side=1, las=1, pos=0, at=x.at, 
       labels=T, xpd=TRUE)
#changing labels=T, puts the ddefault labels back

  #second value should be around .75 (moves label text closer)
  axis(4, las=1, pos=1, at=x.at) #this is the y axis
  lines(0:1, 0:1, xpd=TRUE) #diagonal line
}

@@

  #adjusting this mar lowers the margins (the edges)
  op <- par(mar=c(.25, .15, .25, .75), oma=c(2, 0, 0, 2))
#(1=bottom, 2=left, 3=top, 4=right)

Q2.bottom.triangle2(my.y.bottom, my.x.bottom, my.data) #x.dist

@@

```{r Q2.load.bottom.tri, echo=FALSE}

#only load, SKIVE, PWD, MSM
#MOLF, KAZ, LEW
Q2.bottom.LEW <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_LEW.png')
Q2.bottom.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_KAZ.png')
Q2.bottom.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_SKIVE.png')
Q2.bottom.PWD <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_PWD.png')
Q2.bottom.MSM <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MSM.png')
Q2.bottom.MOLF <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MOLF.png')

#EMPTY GGPLOT
Q2.LEW.bottom.tri <- ggplot(empty.DF, aes(x = col1, y = col2)) +
  background_image(Q2.bottom.LEW) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(),
    axis.title.x=element_blank(),                       axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )

@@

  #drawing triangles in the corner
  #top right-triangle
  draw_line(
    x = c(.1, .2,  .1, .1 ), #keep the x values consistant
    y = c( .25, .25,  .15, .25),
    color = "blue", size=.75
  )

#draw boxes around the

@@

```{r Fig6.IFD.square, echo=FALSE}
#new figure, triangle squares

@@

new figure,
